---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "12-03-2023"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(broom)
library(knitr)
library(kableExtra)
library(tidyverse)
library(dplyr)
```


# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point. 



# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

# Fit the model
model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)

# Get the summary statistics and tidy them up
tidy_summary <- tidy(summary(model))

# Create a data frame for the tidy summary
summary_df <- as.data.frame(tidy_summary)

# # Rename the row names
# row_names <- c("Intercept", "Mother's Prestiges Score", "Father's Prestiges Score", "Father's Education", "Mother's Education")
# rownames(summary_df) <- row_names

# Rename the column names
col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
colnames(summary_df) <- col_names

# Print the summary table with kableExtra
library(kableExtra)
kable(summary_df, align = c("l", "r", "r", "r", "r"), 
      caption = "Summary of Linear Regression Model",) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4
ggplot(data = cleaned_data, aes(x = educ)) +
  
  geom_histogram(aes(y = ..density..), binwidth = 2, fill = "white") +
  geom_density(alpha = .2, fill = "lightblue") +
  
  # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
  labs(title = "Histogram of Respondents' Education", x = "Education (in Years)", y = "Desensity") +
   scale_x_continuous(breaks = seq(0, 20, by = 1))+
  theme_bw()
```

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = mapres10)) +
  geom_histogram(aes(y = ..density.., fill = "mapres10"), alpha = 0.8, binwidth = 2,color = "black") +
  geom_histogram(aes(x = papres10, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 2,color = "black") +
  scale_fill_manual(values = c("mapres10" = "lightblue", "paeduc" = "red"), name = "Variables") +
  labs(title = "Difference in Mother's and Father's Prestige Score",
       x = "Prestige Score",
       y = "Density")+
  theme_bw()
```

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = maeduc)) +
  geom_histogram(aes(y = ..density.., fill = "maeduc"), alpha = 0.8, binwidth = 1, color = "black") +
  geom_histogram(aes(x = paeduc, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 1, color = "black") +
  scale_fill_manual(values = c("maeduc" = "lightblue", "paeduc" = "red"), name = "Variables") +
  labs(title = "Differene in Mother's Education and Father's Education",
       x = "Education (in Years)",
       y = "Density") +
  theme_bw()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(cleaned_data, aes(x = educ, fill = factor(father_factor))) +
  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
   labs(title = "Father's Prestige Score and Education",
       x = "Education (in Years)",
       y = "Count",
       fill = "Level of Prestige Score")+
  theme_bw()
```

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4
ggplot(cleaned_data, aes(x = educ, fill = factor(mother_factor))) +
  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
     labs(title = "Mother's Prestige Score and Education",
       x = "Education (in Years)",
       y = "Count",
       fill = "Level of Prestige Score")+
  theme_bw()
```

Talk way more about it. 



# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.


\newpage


# References


